---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OsteoAnalizer

<!-- badges: start -->
<!-- badges: end -->

The goal of OsteoAnalizer is to clean CSV data sets containing information 
related to Osteoarthritis. The package also aims to analyse 
the relation between the 4 key aspects of Osteoarthritis(i.e fluid levels, 
cartilage thickness, severity levels, and patient age) by output visual elements. 

R Version: R version 4.3.1 (2023-06-16)
Platform: x86_64-apple-darwin20 


## Installation

You can install the development version of OsteoAnalizer like so:

``` r
require("devtools")
devtools::install_github("ika-joshi/OsteoAnalizer", build_vignettes = TRUE)
library("OsteoAnalizer")
```

## Overview

Here is a overview of OsteoAnalizer

``` r
ls("package:OsteoAnalizer")
data(package = "OsteoAnalizer") # optional
browseVignettes("OsteoAnalizer")
```





## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(OsteoAnalizer)
## basic example code
```



You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.


