---
title: "Introduction_OstoAnalizer"
author: "Krutika Joshi"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction_OstoAnalizer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(OsteoAnalizer)
```

## Introduction
**OsteoAnalizer** is a R package with the aim to clean CSV data sets containing information 
related to Osteoarthritis. The function __*cleanCSVFile*__ cleans and organizes 
a CSV file into 4 columns: fluid levels, cartilage thickness, patient age and severity levels. The function __*makeBox*__ visualizes the fluid levels, cartilage thickness and patient age against severity levels using 3 box plots. The function __*makeScatter*__ visualizes the relation between any two variables from fluid levels, cartilage thickness and patient age by using scatter plots. 


**This document gives an overview of the functionalities associated with OsteoAnalizer **. It was written in R Markdown, using the [knitr](https://cran.r-project.org/package=knitr) package for production. 

See `help(package = "OsteoAnalizer")` for further details and references provided by `citation("OsteoAnalizer")`. 

To download **OsteoAnalizer**, use the following commands:
``` r
require("devtools")
devtools::install_github("ika-joshi/OsteoAnalizer", build_vignettes = TRUE)
library("OsteoAnalizer")
```

To list all functions available in the package:
``` r
lsf.str("package:OsteoAnalizer")
```

To list all sample data available in the package:
``` r
data(package = "OsteoAnalizer")
```

<br>

## Quick Tutorial
Use provided mock CSV data to have a quick understanding of the functionalities of this package.

### Step 1. Clean ghe data set
To use the example CSV data:
``` r
# extdata is the folder where the mock data set is present 
# example_data.csv is the name of the data set
testData <- system.file("extdata", "example_data.csv", package = "OsteoAnalizer")
```

To clean the mock CSV dataset:
``` r
result <- CleanCSVFile(testData, "fluid_levels","cartilage_thickness","severity_levels","patient_age")
```

To view the cleaned data set:
``` r
result        # Should contain no NA and should have 4 columns that represent the four variables
```

### Step 2: Bar plot approach
To view distribution of fluid levels, cartilage thickness and patient age against severity levels: 
``` r
box_created <- makeBox(result, "fluid_levels","cartilage_thickness","severity_levels","patient_age")
```
To view the box plots:
``` r
box_created       # Should contain 3 box plots: severity levels vs 3 variables
```

### Step 3.1: Scatter plot approach with line of regression
To view the relation between any two variables from fluid levels, cartilage thickness and patient age
``` r
scatter_created <- makeScatter(result, "fluid_levels","cartilage_thickness")
```
To view the scatter plots:
``` r
scatter_created     # Should contain a scatter plot with the two variables mentioned 
```

### Step 3.2: Scatter plot approach without line of regerssion
To view the relation between any two variables from fluid levels, cartilage thickness and patient age against without the line of regression present
``` r
scatter_created <- makeScatter(result, "fluid_levels","cartilage_thickness", FALSE)
```
To view the scatter plots:
``` r
scatter_created     # Should contain a scatter plot with the two variables mentioned and no line of regression 
```
<br>

## Package References
Joshi, K. (2023) OsteoAnalizer:Analyze data sets related to Osteoarthritis, Unpublished. URL https://github.com/ika-joshi/OsteoAnalizer

<br>
## Other References:
Bates, C. (2023, April 9). Tutorial: Loading and cleaning data with R and
the Tidyverse. Dataquest.
https://www.dataquest.io/blog/load-clean-data-r-tidyverse/

Chatgpt. ChatGPT. (n.d.). https://openai.com/chatgpt

Claus O. Wilke (2017). cowplot: Streamlined Plot Theme and Plot Annotations 
for 'ggplot2'. https://github.com/wilkelab/cowplot.

GeeksforGeeks. (2022, December 26). How to color scatter plot points in
R ?. GeeksforGeeks.
https://www.geeksforgeeks.org/how-to-color-scatter-plot-points-in-r/

Holtz, Y. (n.d.). Control Ggplot2 Boxplot colors. – the R Graph Gallery.
https://r-graph-gallery.com/264-control-ggplot2-boxplot-colors.html

Li, F. (2022) sexDisaggregate:To Obtain and Analyze Sex-disaggregated Datasets,
Unpublished. URL https://github.com/lifangy6/sexDisaggregate"

R box plot. DataMentor. (n.d.). https://www.datamentor.io/r-programming/box-plot

R Core Team (2023). _R: A Language and Environment for Statistical Computing_. 
R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>.

Scatter plot in R programming 🟢 [with examples].
R CODER. (2021, June 18). https://r-coder.com/scatter-plot-r/

Wickham H (2016). ggplot2: Elegant Graphics for Data Analysis. 
Springer-Verlag New York. ISBN 978-3-319-24277-4, https://ggplot2.tidyverse.org.

<br>
```{r}
sessionInfo()
```
