---
title: "Introduction_OstoAnalizer"
author: "Krutika Joshi"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction_OstoAnalizer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(OsteoAnalizer)
```

## Introduction
**OsteoAnalizer** is a R package that aims to clean and analyze CSV files that contain data related to osteoarthritis. The function __*cleanCSVFile*__ cleans and organizes a CSV file into 4 key variables: fluid levels, cartilage thickness, patient age and severity levels. The function __*makeBox*__ provides a box plot visual to aid in osteoarthritis analysis. 
he function __*makeScatter*__ visualizes the relation between any two of the four key variables by using scatter plots. 

**This document provides a overview of OsteoAnalizer's functionalities**
It was written in R Markdown, using the [knitr](https://cran.r-project.org/package=knitr) package for production. 

See reference provided by `citation("OsteoAnalizer")`. 

<br>
## Quick Tutorial
This tutorial will walk through downloading and using **OsteoAnalizer**

### Step 1
To download **OsteoAnalizer**, use the following commands:
```{r}
require("devtools")
devtools::install_github("ika-joshi/OsteoAnalizer", build_vignettes = TRUE)
library("OsteoAnalizer")
```

### Step 2
To examine all functions available in the package type the following in the command:
```{r}
ls("package:OsteoAnalizer")
```

To list all sample data available in OsteoAnalizer type the following in the command line:
```{r}
data(package = "OsteoAnalizer")
```
The output should open a new file called "R data sets" and contain the following information:
```
#Data sets in package ‘OsteoAnalizer’:

OA_data                             
```

### Step 3
This step aims to demonstrates the use of **OsteoAnalizer** to clean a built in mock
data set call **example_data**. This data set is a CSV file which contains data related to
individuals that have osteoarthritis. Specifically, it contains fluid levels,
cartilage thickness, severity levels and the patient_age which are the common 
critical aspects of osteoarthritis research. The data set is raw thus it
contains values such as NA which the **CleanCSVFile** function will clean. 

Type the following to use the **CleanCSVFile** function with the mock data 
set(i.e **example_data**). 
```{r}
# extdata is the folder where example_data is present 
# example_data.csv is the name of the mock data set
# testData stores the path to the example_data.csv file
testData <- system.file("extdata", "example_data.csv", package = "OsteoAnalizer")
# Call the CleanCSVFile function
result <- CleanCSVFile(testData, "fluid_levels","cartilage_thickness","severity_levels","patient_age")
```

To view a snipet of the cleaned data set type the following in the command line argument:
```{r}
# Should contain no NA and should have 4 columns that represent the four variables
result        
```

The console should output
```
fluid_levels cartilage_thickness severity_levels patient_age
1          5.0                 3.0               1           1
2          4.8                 2.9               1           2
4          4.7                 3.0               1           4
5          4.5                 3.5               1           6
6          4.6                 3.0               1           8
7          4.5                 3.5               2          10
```
Note the output displays the top rows of the cleaned dataset, showing no missing
values (NAs) and the four specified variables(i.e fluid_levels, 
cartilage_thickness, severity_levels, patient_age).


### Step 4:
In this step, we will use the cleaned data set and create a bar plot visual. 
To create and view the box plot type the following:
```{r}
# To create the box plots
box_created <- makeBox(result, "fluid_levels","cartilage_thickness","severity_levels","patient_age")
# To view the box plots
box_created
```
The box plots compare the distribution of fluid levels, cartilage thickness, and
patient age across different severity levels. This visualization helps in 
identifying patterns and outliers in the data.

### Step 5a: 
This function will focus on creating a scatter plot to examine the relationship 
between fluid levels, cartilage thickness and patient age. As an example we will 
create a scatter plot to examine the relationship between fluid levels and 
cartilage thickness. 
```{r}
# To create the scatter plot with a line of regression
scatter_created <- makeScatter(result, "fluid_levels","cartilage_thickness")
# To view the scatter plot with a line of regression
scatter_created 
```
The scatter plot illustrates the relationship between fluid levels and cartilage
thickness, with a regression line(in red) indicating the direction of the trend. 
In this way, correlations between these variables can be revealed.


### Step 5b: 
We can also visualize the data without the regression line on the mock data set. 
We will use the same variables mentioned above(i.e fluid levels and cartilage
thickness). 
```{r}
# To create the scatter plot without a line of regression
scatter_created_no_reg <- makeScatter(result, "fluid_levels", "cartilage_thickness", FALSE)
# To view the scatter plot without a line of regression
scatter_created_no_reg
```
This scatter plot is similar to the previous scatter plot, but without the 
regression line. This allows a direct assessment of the distribution of the 
data points.

<br>
## Package References
Joshi, K. (2023) OsteoAnalizer:Analyze data sets related to Osteoarthritis, Unpublished. URL https://github.com/ika-joshi/OsteoAnalizer

<br>
## Other References:
Bates, C. (2023, April 9). Tutorial: Loading and cleaning data with R and
the Tidyverse. Dataquest.
https://www.dataquest.io/blog/load-clean-data-r-tidyverse/

Chatgpt. ChatGPT. (n.d.). https://openai.com/chatgpt

Claus O. Wilke (2017). cowplot: Streamlined Plot Theme and Plot Annotations 
for 'ggplot2'. https://github.com/wilkelab/cowplot.

GeeksforGeeks. (2022, December 26). How to color scatter plot points in
R ?. GeeksforGeeks.
https://www.geeksforgeeks.org/how-to-color-scatter-plot-points-in-r/

Holtz, Y. (n.d.). Control Ggplot2 Boxplot colors. – the R Graph Gallery.
https://r-graph-gallery.com/264-control-ggplot2-boxplot-colors.html

Li, F. (2022) sexDisaggregate:To Obtain and Analyze Sex-disaggregated Datasets,
Unpublished. URL https://github.com/lifangy6/sexDisaggregate"

R box plot. DataMentor. (n.d.). https://www.datamentor.io/r-programming/box-plot

R Core Team (2023). _R: A Language and Environment for Statistical Computing_. 
R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>.

Scatter plot in R programming 🟢 [with examples].
R CODER. (2021, June 18). https://r-coder.com/scatter-plot-r/

Wickham H (2016). ggplot2: Elegant Graphics for Data Analysis. 
Springer-Verlag New York. ISBN 978-3-319-24277-4, https://ggplot2.tidyverse.org.

<br>
```{r}
sessionInfo()
```
