---
title: "Introduction_OsteoAnalizer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction_OsteoAnalizer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.align = "center", 
               out.width = "90%",
               fig.width = 6, fig.height = 5.5,
               dev.args=list(pointsize=10),
               par = TRUE, # needed for setting hook 
               collapse = TRUE, # collapse input & ouput code in chunks
               warning = FALSE)
knit_hooks$set(par = function(before, options, envir)
  { if(before && options$fig.show != "none") 
       par(family = "sans", mar=c(4.1,4.1,1.1,1.1), mgp=c(3,1,0), tcl=-0.5)
})
set.seed(1) # for exact reproducibility
```

## Introduction
**OsteoAnalizer** is an R package designed for the analysis of osteoarthritis data. Specifically, it aims to clean and analyze CSV files that contain data related to osteoarthritis. The diagnosis of osteoarthritis(i.e a common joint disorder) often involves evaluation of a number of parameters such as fluid levels, cartilage thickness, patient's age and severity level. It is helpful to analyze these parameters in order to gain a better understanding of the progression and impact of osteoarthritis. The function __*cleanCSVFile*__ cleans and organizes a CSV file into 4 key variables: fluid levels, cartilage thickness, patient age and severity levels. The function __*makeBox*__ provides a box plot visual to aid in osteoarthritis analysis. The function __*makeScatter*__ visualizes the relation between any two of the four key variables by using scatter plots. 

**This document provides a overview of OsteoAnalizer's functionalities**
It was written in R Markdown, using the [knitr](https://cran.r-project.org/package=knitr) package for production. 

See reference provided by `citation("OsteoAnalizer")`. 

<br>


### Quick Tutorial
This tutorial will walk through downloading and using **OsteoAnalizer**

### Step 1
To download **OsteoAnalizer**, use the following commands. It uses devtools for installation, which ensures that you have the latest version directly from the GitHub repository.
```{r}
require("devtools")
devtools::install_github("ika-joshi/OsteoAnalizer", build_vignettes = TRUE)
library("OsteoAnalizer")
```

### Step 2
To examine all functions available in the package type the following in the command. A list of all the functions available in OsteoAnalizer is displayed in the output, each of which is designed to analyze specific aspects of osteoarthritis data.
```{r}
ls("package:OsteoAnalizer")
```

To list all the sample data available in OsteoAnalizer type the following in the command line. **OA_data** is the sample dataset included which provides a realistic representation of clinical data used in osteoarthritis research. The data set has 4 columns with each representing a variable(i.e fluid levels, cartilage thickness, patient's age and severity level). The rows represent the individuals that were tested for the data. 
```{r}
data(package = "OsteoAnalizer")
```
The output should open a new file called "R data sets" and contain the following information:
```
#Data sets in package ‘OsteoAnalizer’:

OA_data                             
```

### Step 3
This step aims to demonstrates the use of **OsteoAnalizer** to clean a built in mock
data set call **example_data**. This data set is a CSV file which contains data related to
individuals that have osteoarthritis. Specifically, it contains fluid levels,
cartilage thickness, severity levels and the patient_age which are the common 
critical aspects of osteoarthritis research. The data set is raw thus it
contains values such as NA which the **CleanCSVFile** function will clean. 

Type the following to use the **CleanCSVFile** function with the mock CSV data 
set(i.e **example_data**). 
```{r}
# extdata is the folder where example_data is present 
# example_data.csv is the name of the mock data set
# testData stores the path to the example_data.csv file
testData <- system.file("extdata", "example_data.csv", package = "OsteoAnalizer")
# Call the CleanCSVFile function
result <- CleanCSVFile(testData, "fluid_levels","cartilage_thickness","severity_levels","patient_age")
```

To view a snipet of the cleaned data set type the following in the command line argument:
```{r}
# Should contain no NA and should have 4 columns that represent the four variables
result        
```
Note the output displays the top rows of the cleaned dataset, showing no missing
values (NAs) and the four specified variables(i.e fluid_levels, 
cartilage_thickness, severity_levels, patient_age). To ensure the integrity and 
reliability of future analyses, it is crucial to thoroughly clean the dataset 
before applying any computational techniques.


### Step 4:
In this step, we will use the cleaned data set and create a bar plot visual. 
To create and view the box plot type the following:
```{r}
# To create the box plots
box_created <- makeBox(result, "fluid_levels","cartilage_thickness","severity_levels","patient_age")
# To view the box plots
box_created
```
This set of box plots illustrates key clinical measurements across four severity 
levels of osteoarthritis. As severity increases, both fluid levels and cartilage
thickness decrease, with the median fluid level being highest in the least severe 
category and lowest in the most severe. Conversely, there is a positive correlation 
between patient age and severity, indicating that older patients are likely to 
experience more advanced stages of the disease. As severity levels increase, the 
interquartile ranges for fluid levels and cartilage thickness widen, suggesting a 
greater degree of variability in the disease's progression. As a result of these
visualizations, we are able to demonstrate relationships between clinical 
measurements and the severity of osteoarthritis, which can be used to formulate 
further research and patient management strategies.

### Step 5a: 
**makeScatter** function will focus on creating a scatter plot to examine the relationship 
between any two of the following variables: fluid levels, cartilage thickness and patient age. As an example, we will 
create a scatter plot to examine the relationship between fluid levels and 
cartilage thickness. 
```{r}
# To create the scatter plot with a line of regression
scatter_created <- makeScatter(result, "fluid_levels","cartilage_thickness")
# To view the scatter plot with a line of regression
scatter_created 
```
According to the scatter plot presented, fluid levels are correlated with 
cartilage thickness in osteoarthritis. This equation suggests that for every
unit increase in fluid levels, cartilage thickness increases by approximately 
0.21 units, as indicated by the upward trend of the regression line. Observing 
this pattern may indicate that higher fluid levels are associated with thicker 
cartilage, which could be significant in the progression or treatment of the 
disease. Based on the spread of data points around the line, it appears that 
this relationship varies from person to person. This visual analysis provides 
a clear depiction of how two critical clinical measurements may interact, 
offering valuable insights for medical research and patient care.


### Step 5b: 
We can also visualize the data without the regression line on the mock data set. 
We will use the same variables mentioned above(i.e fluid levels and cartilage
thickness). 
```{r}
# To create the scatter plot without a line of regression
scatter_created_no_reg <- makeScatter(result, "fluid_levels", "cartilage_thickness", FALSE)
# To view the scatter plot without a line of regression
scatter_created_no_reg
```
This scatter plot is similar to the previous scatter plot, but without the 
regression line. This allows a direct assessment of the distribution of the 
data points. In this plot, the data points are dispersed across the plot. The 
dispersion can imply a more complex relationship that may  
be influenced by other factors not represented in this plot. This might suggest 
researchers to consider additional variables or more sophisticated models to 
understand the dynamics between fluid levels and cartridge thickness.



<br>
## Package References
Joshi, K. (2023) OsteoAnalizer:Analyze data sets related to Osteoarthritis, Unpublished. URL https://github.com/ika-joshi/OsteoAnalizer

<br>
## Other References:
Bates, C. (2023, April 9). Tutorial: Loading and cleaning data with R and
the Tidyverse. Dataquest.
https://www.dataquest.io/blog/load-clean-data-r-tidyverse/

Chatgpt. ChatGPT. (n.d.). https://openai.com/chatgpt

Claus O. Wilke (2017). cowplot: Streamlined Plot Theme and Plot Annotations 
for 'ggplot2'. https://github.com/wilkelab/cowplot.

GeeksforGeeks. (2022, December 26). How to color scatter plot points in
R ?. GeeksforGeeks.
https://www.geeksforgeeks.org/how-to-color-scatter-plot-points-in-r/

Holtz, Y. (n.d.). Control Ggplot2 Boxplot colors. – the R Graph Gallery.
https://r-graph-gallery.com/264-control-ggplot2-boxplot-colors.html

Li, F. (2022) sexDisaggregate:To Obtain and Analyze Sex-disaggregated Datasets,
Unpublished. URL https://github.com/lifangy6/sexDisaggregate"

R box plot. DataMentor. (n.d.). https://www.datamentor.io/r-programming/box-plot

R Core Team (2023). _R: A Language and Environment for Statistical Computing_. 
R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>.

Scatter plot in R programming 🟢 [with examples].
R CODER. (2021, June 18). https://r-coder.com/scatter-plot-r/

Wickham H (2016). ggplot2: Elegant Graphics for Data Analysis. 
Springer-Verlag New York. ISBN 978-3-319-24277-4, https://ggplot2.tidyverse.org.

<br>
```{r}
sessionInfo()
```
